---
title: "Quant Assignment 2"
author: "Marissa Getts"
date: "9/16/2020"
output: html_document
---
```{r}
hh_data <- read.csv("MGquant1Final.csv")
```

```{r}
library(tidyverse)
library(ggplot2)
attach(hh_data)
```

Here's a quick overview of my variables in the above .csv:  
<i>Categorical:</i>  <ol><li>RACAIAN - American Indian/Alaska Native recode
             <li> ACCESS - access to internet
             <li> LAPTOP - laptop or desktop
             <li> TABLET - tablet or other computer
             <li> COMPOTHX - other computer equipment
             <li> SMARTPHONE - smartphone
             <li> BROADBND - cellular data plan for a smartphone or other mobile device
             <li> HUPAC - age and presence of children under 18</ol>
              
<i>Continuous:</i>  <ol><li> HINCP - household income past 12 months
             <li> BDSP - bedrooms per household
             <li>NP - number of persons per household
             <li> NOC - number of own children

<b>Summaries</b></br>

The code and summaries - including mean, median, interquartile range, standard deviation, and confidence intervals are included below. <br>

<i>Household Income Summary</i>
```{r HINCP}
HINCP_summary <- summary(hh_data$HINCP)
HINCP_summary["Median"]
quantile(hh_data$HINCP)
sd(hh_data$HINCP)
conf_int <- t.test(hh_data$HINCP)
conf_int
```
<br><i>Bedrooms per Household Summary</i>
```{r BDSP}

BDSP_summary <- summary(hh_data$BDSP)
BDSP_summary["Median"]
quantile(hh_data$BDSP)
sd(hh_data$BDSP)
conf_int <- t.test(hh_data$BDSP)
conf_int
```
<br><i>Persons per Household Summary</i>
```{r NP}
NP_summary <- summary(hh_data$NP)
NP_summary["Median"]
quantile(hh_data$NP)
sd(hh_data$NP)
conf_int <- t.test(hh_data$NP)
conf_int
```
<br><i>Own Children per Household</i>
```{r NOC}
NOC_summary <- summary(hh_data$NOC)
NOC_summary["Median"]
quantile(hh_data$NOC)
sd(hh_data$NOC)
conf_int <- t.test(hh_data$NOC)
conf_int
```
<br><i>Related Children per Household</i>
```{r NRC}
NRC_summary <- summary(hh_data$NRC)
NRC_summary["Median"]
quantile(hh_data$NRC)
sd(hh_data$NRC)
conf_int <- t.test(hh_data$NRC)
conf_int
```


<b>HISTOGRAMS</b><br>
Here are the histograms for each of my continuous variables. 

<br><i>Histogram: Household Income</i>
```{r}
hist1 <- ggplot(hh_data, aes(x = HINCP)) +
  geom_histogram() 
```

```{r}
hist1+
  theme_bw() +
  scale_x_continuous(name = "Household Income in USD") +
  scale_y_continuous(name = "Total") +
  annotate("text", x = 500000, y = 8000, size = 3, 
           label = paste("\nS mean:", 
                         prettyNum(mean(HINCP), digits = 0), 
                         "\nS median:",
                         prettyNum(median(HINCP), digits = 0),
                         "\nS standard deviation:",
                         prettyNum(sd(HINCP), digits = 0), 
                         "\nS interquartile range:",
                         prettyNum(quantile(HINCP)["25%"], digits = 0),
                         "to",
                         prettyNum(quantile(HINCP)["75%"], digits = 0)),
           hjust = 0, vjust = 1) 
```
<br><i>Histogram: Bedrooms per Household</i>

```{r}
hist2 <- ggplot(hh_data, aes(x = BDSP)) +
  geom_histogram() 
```

```{r}
hist2+
  theme_bw() +
  scale_x_continuous(name = "Bedrooms per Household") +
  scale_y_continuous(name = "Frequency") +
  annotate("text", x = 6, y = 6000, size = 3, 
           label = paste("\nSample mean:", 
                         prettyNum(mean(BDSP), digits = 0), 
                         "\nSample median:",
                         prettyNum(median(BDSP), digits = 0),
                         "\nSample standard deviation:",
                         prettyNum(sd(BDSP), digits = 0), 
                         "\nSample interquartile range:",
                         prettyNum(quantile(BDSP)["25%"], digits = 0),
                         "to",
                         prettyNum(quantile(BDSP)["75%"], digits = 0)),
           hjust = 0, vjust = 1) 
```
<br><i>Histogram: People per Household</i>
```{r}
hist3 <- ggplot(hh_data, aes(x = NP)) +
  geom_histogram()
```

```{r}
hist2+
  theme_bw() +
  scale_x_continuous(name = "People per Household") +
  scale_y_continuous(name = "Frequency") +
  annotate("text", x = 6, y = 7000, size = 3, 
           label = paste( "\nSample mean:", 
                         prettyNum(mean(NP), digits = 0), 
                         "\nSample median:",
                         prettyNum(median(NP), digits = 0),
                         "\nSample standard deviation:",
                         prettyNum(sd(NP), digits = 0), 
                         "\nSample interquartile range:",
                         prettyNum(quantile(NP)["25%"], digits = 0),
                         "to",
                         prettyNum(quantile(NP)["75%"], digits = 0)),
           hjust = 0, vjust = 1) 
```


<br><i>Histogram:Own Children per Household</i>

```{r}
hist4 <- ggplot(hh_data, aes(x = NOC)) +
  geom_histogram()
```

```{r}
hist2+
  theme_bw() +
  scale_x_continuous(name = "Own Children Per Household") +
  scale_y_continuous(name = "Frequency") +
  annotate("text", x = 6, y = 6000, size = 3, 
           label = paste( "\nSample mean:", 
                         prettyNum(mean(NOC), digits = 0), 
                         "\nSample median:",
                         prettyNum(median(NOC), digits = 0),
                         "\nSample standard deviation:",
                         prettyNum(sd(NOC), digits = 0), 
                         "\nSample interquartile range:",
                         prettyNum(quantile(NOC)["25%"], digits = 0),
                         "to",
                         prettyNum(quantile(NOC)["75%"], digits = 0)),
           hjust = 0, vjust = 1) 
```

<br><i>Histogram:Related Children per Household</i>
```{r}
hist5 <- ggplot(hh_data, aes(x = NRC)) +
  geom_histogram()
```

```{r}
hist2+
  theme_bw() +
  scale_x_continuous(name = "Related Children Per Household") +
  scale_y_continuous(name = "Frequency") +
  annotate("text", x = 6, y = 6000, size = 3, 
           label = paste( "\nSample mean:", 
                         prettyNum(mean(NRC), digits = 0), 
                         "\nSample median:",
                         prettyNum(median(NRC), digits = 0),
                         "\nSample standard deviation:",
                         prettyNum(sd(NRC), digits = 0), 
                         "\nSample interquartile range:",
                         prettyNum(quantile(NRC)["25%"], digits = 0),
                         "to",
                         prettyNum(quantile(NRC)["75%"], digits = 0)),
           hjust = 0, vjust = 1) 
```

