---
title: "Quant 4"
author: "Marissa Getts"
date: "9/29/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(ggplot2)
library(sf)
library(rnaturalearth)
library(tidyverse)
library(tidycensus)
library(ggspatial)
library(cartogram)
library(ggthemes)
library(dplyr)
library(dbplyr)
library(ggthemr)
```

```{r looking through household variables, include=FALSE}
hh_vars_2018 <- pums_variables %>%
distinct(year, survey, var_code, 
           var_label, data_type, level) %>%
filter(level == "household", year == 2018, survey == "acs1")
```


```{r load household level data, include=FALSE}
hh_data <- get_pums(variables = c("ACCESS",
                                  "LAPTOP",
                                  "SMARTPHONE",
                                  "HINCP", 
                                  "BDSP",
                                  "NP",
                                  "NOC"),
                        state = "UT",
                        year = 2018, 
                        survey = "acs1",
                        recode = TRUE) %>%
  filter(HINCP > 0, NOC > -1, BDSP > -1)
```

### producing regression results

```{r}
model <- lm(HINCP ~ BDSP + NP + NOC + SMARTPHONE + LAPTOP + ACCESS, 
            data = hh_data)

options(scipen=999)

summary(model)
```
### how well does the model fit the data?

The multiple R squared value for this data set is .1505, meaning that this model predicts about 15% of the variation in income for the households in this dataset. Although I don't have much to compare it to, it is much more substantial than the example in the tutorial (4%). So, while it is nothing like 100%, it is likely a higher fit than many regression models (especially in the social sciences) may find. 

### explanation of variable coefficients + significance

<ol><li><b>income and number of bedrooms:</b> the coefficient for BDSP is 21469.3, meaning that for each additional bedroom, a household's income is expected to be $21,469 greater, controlling for number of people per household, number of own children per household, smartphone/laptop ownership, and access to internet. The p value is statistically significant. 

<li><b>income and number of people</b>: the coefficient for NP is 7949.3, meaning that for each addition person, a household's income is expected to be $7949.30 greater, controlling for bedrooms per household, number of own children, smartphone/laptop ownership, and access to internet. The p value is statistically significant.

<li><b>income and number of own children:</b> the coefficient for NOC is -10709.1, meaning that for each additional own child, a household's income is expected to be $10,709.10 less, controlling for bedrooms per household, number of people per household, smartphone/laptop ownership, and access to internet. The p value is statistically significant.

<li><b>income and smartphone ownership:</b> the coefficient for SMARTPHONE2 is -18411.8, meaning that, on average and controlling for all other variables in this set, if a household does not own a smartphone, the household's income is expected to $18,411.80 lower than a household that does own a smartphone. The p value is statistically significant. 

<li><b>income and laptop ownership:</b> the coefficient for LAPTOP2 is -19296.7, meaning that, on average and controlling for all other variables in the set, if a household does not own a laptop, the household's income is expected to be $19,296.70 lower than the income of a household that does own a laptop. The p value is statistically significant. 

<li><b>income and internet access:</b> this particular set had multiple variables, so it is comparing them both against households that have internet access by "paying a cell phone company or internet service provider." the coefficient for ACCESS2 is -12800.4, meaning that, on average and controlling for all other variables in the set, if a household does not have access to the internet, the household's income is expected to be $12,800.40 lower than the income of a household that has internet access by paying a cell phone company or internet service provider. the coefficient for ACCESS3 is -1291.4, meaning that (if it was statistically significant) a household in this data set that has access to the internet *without* paying a cell phone company or internet service provider would be associated with an income $1291.4 lower than a household that has internet access by paying a cell phone company or internet service provider. ACCESS3, however, has a p value of .683, meaning that it is not statistically significant, and therefore predicts nothing. :) 

</ol>
